Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2020-09-03 Manuscript/HI_Model_ms.Rmd
updated manuscript files

## Introduction

The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes,

## Methods
We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands under all combinations of two land-use scenarios (low and high) and two radiative forcing scenarios (RCP 4.5 and RCP 8.5). The landscape change portion of LUCAS is a state-and-transition model that applies a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transitions [@daniel_state-and-transition_2016]. The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying stock level rates of change over time [@daniel_integrating_2018; @sleeter_effects_2019]. We parameterized the Hawai`i LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability. Simulations were run for 90 years at an annual timestep, using initial conditions corresponding to the year 2010. All simulations were repeated for 30 Monte Carlo realizations.

@@

### *CO~2~ fertilization effect*

Increasing atmospheric CO~2~ concentrations stimulate leaf-level photosynthesis, potentially increasing NPP as well. However, the magnitude and persistence of this effect is highly uncertain, particularly across a range of climatic conditions and over long time spans. Following @sleeter_effects_2019, we developed a separate set of scenarios designed to test the sensitivity of LUCAS model ecosystem carbon balance projections to the influence of differing rates of a CO~2~ fertilization effect (CFE). We incorporated a NPP CFE multiplier representing the percent increase in NPP for every 100 ppm increase in atmospheric CO~2~ concentration under the high land use and high radiative forcing (RCP 8.5) scenario. We tested five CFE levels ranging from 5% to 15%, which is within the range of CFEs observed in free air CO~2~ enrichment (FACE) experiments. For all levels, we assumed CFEs reached saturation at an atmospheric CO~2~ concentration of 600 ppm, with no further increase in NPP despite a continued increase in CO~2~ concentration to 930 ppm by 2100. This 600ppm threshold generally coincides with the upper limit from FACE experiments and is reached by the year 2060 under RCP 8.5.

### *Scenario simulations and analysis*

## Results

>>>>

## Introduction
The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes,

## Methods
We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands under all combinations of two land-use scenarios (low and high) and two radiative forcing scenarios (RCP 4.5 and RCP 8.5). We also developed a separate set of scenarios to test model sensitivity to different levels of a CO~2~ fertilization effect (CFE). The landscape change portion of LUCAS is a state-and-transition model that applies a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transitions [@daniel_state-and-transition_2016]. The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying stock level rates of change over time [@daniel_integrating_2018; @sleeter_effects_2019]. We parameterized the Hawai`i LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability for the time period 2010-2100.

@@

### *CO~2~ fertilization effect*
Increasing atmospheric CO~2~ concentrations stimulate leaf-level photosynthesis, potentially increasing NPP as well. However, the magnitude and persistence of this effect is highly uncertain, particularly across a range of climatic conditions and over long time spans. Following @sleeter_effects_2019, we developed a separate set of scenarios designed to test the sensitivity of LUCAS model ecosystem carbon balance projections to the influence of differing rates of a CO~2~ fertilization effect (CFE). We incorporated a NPP CFE multiplier representing the percent increase in NPP for every 100 ppm increase in atmospheric CO~2~ concentration under the high land use and high radiative forcing (RCP 8.5) scenario. We tested five CFE levels ranging from 5% to 15%, which is within the range of CFEs observed in free air CO~2~ enrichment (FACE) experiments. For all levels, we assumed CFEs reached saturation at an atmospheric CO~2~ concentration of 600 ppm, with no further increase in NPP despite a continued increase in CO~2~ concentration to 930 ppm by 2100. This 600ppm threshold generally coincides with the upper limit from FACE experiments and is reached by the year 2060 under RCP 8.5.

### *Scenario simulations and analysis*
Each of the four unique scenarios were run for 90 years at an annual timestep and repeated for 30 Monte Carlo realizations, using initial conditions corresponding to the year 2010. All simulations were performed within the SyncroSim (version 2.2.4) software framework with ST-Sim (version 3.2.13) and SF (version 3.2.10) add-on modules [@daniel_state-and-transition_2016; @daniel_integrating_2018]. Model inputs and outputs were prepared with the R statistical computing platform [@RCore_2019] using the tidyverse [@Wickham_tidyverse_2019], raster [@hijmans_raster_2020], and rsyncrosim [@daniel_rsyncrosim_2020] packages.

## Results
 Terrestrial ecosystems of the seven main Hawaiian Islands stored an estimated 316 Tg of carbon at the beginning of the simulation period in 2010, with 58% in soil organic matter, 22% in living biomass, and 20% in surface dead organic matter (litter and dead wood; Fig. 2a). The combination of low radiative forcing and low land use change yielded the highest and most consistent projections of ecosystem carbon accumulation, with a gain of ~15% to an average of 363 Tg of carbon by 2100 (Fig. 2a).

```{r fig.align = 'center', out.width = "95%", fig.cap = "Projected changes in total ecosystem carbon storage (a) and individual carbon stocks (b) for the seven main Hawaiian Islands. Solid lines indicate the mean of 30 Monte Carlo realizations for each scenario, with shaded areas indicating the minimum and maximum range of Monte Carlo realizations."}
knitr::include_graphics("./fig_images/fig2_Cstocks.png")
```
